---
title: "Lumpfish Visuals"
author: "Sophie Wulfing"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
---
This is a visualization of Lumpfish data after the latest NOAA Observer data was incorporated

NOAA Observer Notes:

 * NOAA Observer data was not uploaded to GitHub due to confidentiality. Must be on local computer to run analyses
 
 * In NOAA Observer Data, only the samples that were lengthed have number of fish per tow specified. Otherwise, we just know that at least one fish was caught on this tow, making the WEIGHT_TOW column useless in non-lengthed observations
 
 * When filtering for fish below 61 cm and 9.5 kg, only NOAA Observer data was eliminated. All other datasets were inside that range
 
 * TOW IDs in NOAA Observer data were eliminated due to confidentiality
 
 * Some weights were confusing (See boxplot below)


```{r setup, include=FALSE, cache=FALSE, echo = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(sf)
library(mapview)
library(lubridate)
library(gganimate)
library(gifski)
library(maps)
library(knitr)
library(ggpubr)
library(magick)
library(ggplot2)
library(dplyr)

setwd("C:/Users/sophi/Documents/GitHub/TeamLump")
Lumpfish_Data_clean <- read.csv("DatatoClean/Lumpfish_Data_clean.csv")

Lumpfish_Data_clean$YEAR <- year(Lumpfish_Data_clean$DATE)

Lumpfish_Data_map <- Lumpfish_Data_clean %>%
  drop_na(c("LAT", "LON", "DATE", "YEAR")) %>%
  mutate(across(SEASON, factor, levels = c("WINTER", "SPRING", "SUMMER", "FALL")))

which_state <- c("maine", "new hampshire", "vermont", "massachusetts", "rhode island", "connecticut",
                 "new york", "new jersey", "pennsylvania", "virginia", "maryland", "washington dc", "delaware")
county_info <- map_data("state", region=which_state)

base_map <- ggplot(data = county_info, mapping = aes(x = long, y = lat, group = group)) +
  geom_polygon(color = "black", fill = "white") +
  coord_quickmap() +
  theme_void() 

min_long <- min(Lumpfish_Data_map$LON)
max_long <- max(Lumpfish_Data_map$LON)
min_lat <- 41.5 #min(Lumpfish_Data_map$LAT)
max_lat <- max(Lumpfish_Data_map$LAT)

zoom_map <- base_map +
   coord_quickmap(xlim = c(min_long, max_long),  ylim = c(min_lat, max_lat))


# testing <- Lumpfish_Data_clean %>%
#   filter(DATA_SOURCE == "NEFSC") %>%
#   #drop_na(c("LAT", "LON", "DATE", "YEAR")) %>%
#   filter(is.na(YEAR))
# 
# 
# View(testing)
# 
# 
# min(testing$YEAR)
```

```{r pointsperyear, echo = FALSE, fig.cap='Datapoints per year. Totals 12,920 observations spanning from 1963 to 2023', warning=FALSE}

#sorted_years <- lapply(list(unique(Lumpfish_Data_clean$YEAR)), sort)
library(ggplot2)

yearMap_zoom <- zoom_map +
  geom_point(data = Lumpfish_Data_map, aes(x = LON, y = LAT, group=YEAR, colour = YEAR))
yearMap_zoom

```

\newpage

```{r pointstable, echo = FALSE, results = 'asis', warning=FALSE}

library(dplyr)

year_table <- data.frame(matrix(ncol = 3, nrow = 0))
colnames(year_table) <- c("Data Code", "Years Spanned", "Number of datapoints")

sources <- unique(Lumpfish_Data_clean$DATA_SOURCE)

for(i in 1:length(sources)){
  new_data <- Lumpfish_Data_clean %>%
    filter(DATA_SOURCE == sources[i])
  
  year_table[i,] <- c(sources[i], paste0(min(new_data$YEAR), "-", max(new_data$YEAR)), nrow(new_data))
}

knitr::kable(year_table, caption = "Data points and year spanned per source")

```

\newpage

```{r FishPerSEASON, echo=FALSE, fig.cap='Datapoints per season', warning=FALSE}
library(dplyr)

#FIX LEGEND
season_labs <- c("Winter", "Spring", "Summer", "Fall")
names(season_labs) <- c("WINTER", "SPRING", "SUMMER", "FALL")

zoom_map +
  geom_point(data = Lumpfish_Data_map, aes(x = LON, y = LAT, group=DATA_SOURCE, colour = DATA_SOURCE), size = .75) +
  facet_wrap(~ SEASON, ncol = 2, labeller = labeller(SEASON = season_labs))



```

```{r FishPerSeasonAnimation, echo = FALSE, fig.cap='Datapoints per season', warning=FALSE}

library(gganimate)
map_with_seasondata <- zoom_map +
  geom_point(data = Lumpfish_Data_map, aes(x = LON, y = LAT, group=SEASON, colour = DATA_SOURCE), size = 0.75)

map_with_seasonanimation <- map_with_seasondata +
  transition_states(SEASON) +
  ggtitle('Season: {next_state}')
#num_years <- max(Lumpfish_Data_map$YEAR) - min(Lumpfish_Data_map$YEAR) + 1
animate(map_with_seasonanimation, fps = 1, nframes = 5)

```

\newpage

```{r LenperYear, echo = FALSE, fig.cap = "Average Length of lump fish over time", warning=FALSE}
library(tidyr)
library(lubridate)

yearly_length <- Lumpfish_Data_clean %>%
  mutate(YEAR = year(DATE)) %>%
  drop_na(LENGTH) %>%
  group_by(YEAR) %>%
  summarize(LENGTH_MEAN = mean(LENGTH))

plot(yearly_length, ylab="Mean Length (cm)", xlab="Year")
```

\newpage

```{r WtperYear, echo = FALSE, fig.cap = "Average weight of fish over time across all data sources", warning=FALSE}
yearly_wt <- Lumpfish_Data_clean %>%
  mutate(YEAR = year(DATE)) %>%
  drop_na(WEIGHT_AVG) %>%
  group_by(YEAR) %>%
  summarize(WT_MEAN = mean(WEIGHT_AVG))

plot(yearly_wt, ylab="Mean weight (kg)", xlab="Year")
```

```{r WtperYearAll, echo = FALSE, fig.show="hold", out.width="50%", results = "asis", fig.cap = "Average weight of fish over time with outlier datasets (MA_DMF and NEFSC) and the newer dataset (NOAA_Observer) to show how they're skewing data", warning=FALSE}
yearly_wt <- Lumpfish_Data_clean %>%
  #mutate(YEAR = year(DATE)) %>%
  drop_na(WEIGHT_AVG) %>%
  group_by(YEAR) %>%
  summarize(WT_MEAN = mean(WEIGHT_AVG))

plot(yearly_wt, main = "Average weight of catch across all data sources", ylab="Mean weight (kg)", xlab="Year")


MA_DMF <- Lumpfish_Data_clean %>%
  filter(DATA_SOURCE == "MA_DMF") %>%
  #mutate(YEAR = year(DATE)) %>%
  drop_na(WEIGHT_AVG) %>%
  group_by(YEAR) %>%
  summarize(WT_MEAN = mean(WEIGHT_AVG))

plot(MA_DMF, main = "MA_DMF average weight of catch over time", ylab="Mean weight (kg)", xlab="Year")

NEFSC <- Lumpfish_Data_clean %>%
  filter(DATA_SOURCE == "NEFSC") %>%
  mutate(YEAR = year(DATE)) %>%
  drop_na(WEIGHT_AVG) %>%
  group_by(YEAR) %>%
  summarize(WT_MEAN = mean(WEIGHT_AVG))

plot(NEFSC, main = "NEFSC average weight of catch over time", ylab="Mean weight (kg)", xlab="Year")

NOAA_Observer <- Lumpfish_Data_clean %>%
  filter(DATA_SOURCE == "NOAA_Observer") %>%
  mutate(YEAR = year(DATE)) %>%
  drop_na(WEIGHT_AVG) %>%
  group_by(YEAR) %>%
  summarize(WT_MEAN = mean(WEIGHT_AVG))

plot(NOAA_Observer, main = "NOAA_Observer average weight of catch over time", ylab="Mean weight (kg)", xlab="Year")
```


```{r avgweightBoxplot, echo = FALSE, fig.cap = "Range of reported average weights (kg) per datasource. Average weights were taken when weight of whole tow was reported as well as number of fish in the tow", warning=FALSE}

ggplot(Lumpfish_Data_map, aes(x = DATA_SOURCE, y = WEIGHT_AVG)) +
  geom_boxplot()


```

```{r weightBoxplot, echo = FALSE, fig.cap = "Range of reported individual weights (kg) per datasource. Individual weights were used when data collectors individually weight each fish", warning=FALSE}

ggplot(Lumpfish_Data_map, aes(x = DATA_SOURCE, y = WEIGHT)) +
  geom_boxplot()


```

```{r lengthBoxplot, echo = FALSE, fig.cap = "Range of reported individual weights (kg) per datasource. Individual weights were used when data collectors individually weight each fish", warning=FALSE}

ggplot(Lumpfish_Data_map, aes(x = DATA_SOURCE, y = LENGTH)) +
  geom_boxplot()


```

\newpage

```{r animationCentroid, echo = FALSE, fig.cap = "Centroid of lumpfish found per year (left) with raw data per year (right)", warning=FALSE}

library(magick)

Lumpfish_With_Centroid <- Lumpfish_Data_map %>%
  group_by(YEAR) %>%
  mutate(LAT.AVG = mean(LAT)) %>%
  mutate(LON.AVG = mean(LON))


map_with_rawdata <- zoom_map +
  geom_point(data = Lumpfish_With_Centroid, aes(x = LON, y = LAT, group=YEAR))

map_with_centroiddata <- zoom_map +
  geom_point(data = Lumpfish_With_Centroid, aes(x = LON.AVG, y = LAT.AVG, group=YEAR))

# require(gridExtra)
# map_panel <- grid.arrange(map_with_centroiddata, map_with_rawdata, ncol=2)

map_with_raw_animation <- map_with_rawdata +
  transition_time(YEAR) +
  ggtitle('Year: {frame_time}',
          subtitle = 'Frame {frame} of {nframes}')
num_years <- max(Lumpfish_Data_map$YEAR) - min(Lumpfish_Data_map$YEAR) + 1
b_gif <- animate(map_with_raw_animation, nframes = num_years, fps = 2)

map_with_centroid_animation <- map_with_centroiddata +
  transition_time(YEAR) +
  ggtitle('Year: {frame_time}',
          subtitle = 'Frame {frame} of {nframes}')
num_years <- max(Lumpfish_Data_map$YEAR) - min(Lumpfish_Data_map$YEAR) + 1
a_gif <- animate(map_with_centroid_animation, nframes = num_years, fps = 2)

a_mgif <- image_read(a_gif)
b_mgif <- image_read(b_gif)

new_gif <- image_append(c(a_mgif[1], b_mgif[1]))
for(i in 2:60){
  combined <- image_append(c(a_mgif[i], b_mgif[i]))
  new_gif <- c(new_gif, combined)
}

new_gif
```

# Age and Bayesian Weights


```{r BayesweightBoxplot, echo = FALSE, fig.cap = "Range of calculated Bayesian weights per datasource"}

bayes_graph <- Lumpfish_Data_map %>%
  filter(!is.na(WEIGHT_BAYESIAN))

ggplot(bayes_graph, aes(x = DATA_SOURCE, y = WEIGHT_BAYESIAN)) +
  geom_boxplot()


```

```{r agedistBarplot, echo = FALSE, fig.cap = "Age of fish caught by each Data Source" }


yearlyStacked <- Lumpfish_Data_map %>%
  group_by(DATA_SOURCE, AGE) %>%
  filter(!is.na(AGE)) %>%
  summarize(count = n())

ggplot(yearlyStacked, aes(fill=AGE, y=count, x=DATA_SOURCE)) + 
  geom_bar(position='stack', stat='identity')

#Then do age over years
#Age over season

```

```{r ageperyearBarplot, echo = FALSE, fig.cap = "Number of datapoints per year separated by age" }


yearlyageStacked <- Lumpfish_Data_map %>%
  group_by(YEAR, AGE) %>%
  filter(!is.na(AGE)) %>%
  summarize(count = n())

ggplot(yearlyageStacked, aes(fill=AGE, y=count, x=YEAR)) + 
  geom_bar(position='stack', stat='identity')


```

```{r ageperseasonBarplot, echo = FALSE, fig.cap = "Number of datapoints per season separated by age" }


seasonageStacked <- Lumpfish_Data_map %>%
  group_by(SEASON, AGE) %>%
  filter(!is.na(AGE)) %>%
  summarize(count = n())

ggplot(seasonageStacked, aes(fill=AGE, y=count, x=SEASON)) + 
  geom_bar(position='stack', stat='identity')


```






